\subsection{Unplanned Automatic Scrams Per 7,000 Hours Critical (UA7)}

\subsubsection{Purpose}

The purpose of the unplanned automatic scrams per 7,000 hours critical
indicator is to monitor performance in reducing the number of
unplanned automatic reactor shutdowns. The indicator provides an
indication of success in improving plant safety by reducing the number
of undesirable and unplanned thermal-hydraulic and reactivity
transients that require reactor scrams. It also provides an indication
of how well a plant is operated and maintained. 

Taking into account the number of hours that a plant is critical
provides an indication of the effectiveness of scram reduction efforts
while a unit is in an operating condition. In addition, normalising
individual unit scram data to a common standard (7,000 hours critical)
provides a uniform basis for comparisons among individual units and
with industry values. 

\subsubsection{Definition}

The indicator is defined as the number of unplanned automatic scrams
(reactor protection system logic actuations) that occur per 7,000
hours of critical operation.
 
The value of 7,000 hours is representative of the critical hours of
operation during a year for most plants. It provides an indicator
value that typically approximates the actual number of scrams
occurring during the year.

\subsubsection{Calculations}

The unit and industry values for this indicator are determined for a period as shown below:  
$$ Value for a unit = \frac{AS \cdot 7000}{total number of hours critical} $$	
Where 	AS = number of unplanned automatic scrams

$$ Value for the industry = mean of the unit values $$

Because these calculations are based on the number of scrams resulting
per 7,000 critical hours, the typical result for both an individual
unit or worldwide will not be an integer.  For comparisons of
individual units (e.g. histograms) unplanned automatic scrams per
7,000 hours critical will be presented for a three-year period to
minimise the effects of variations in the indicator value during
shorter time periods due to the low number of scrams at most
plants. Units must average at least 1,000 critical hours per year to
be calculated and included in the worldwide value.

An example of the data collected for this performance indicator and a
sample calculation are provided in the Attachment \ref{UA7}.

